Approximate Bayesian Fitting

### 2023-05-23 Misc/benchmarks.qmd
more image changes and misc folder

---
title: Benchmarking 
code-fold: show
code-tools: true
execute: 
  warning: false
  eval: false
---

Furrr and standard version both work equally fast, tested with up to 120 simulation repetitions
```{r}
library(furrr)
furrr::furrr_options(seed = TRUE)
plan(multisession, workers = parallel::detectCores())

@@

d <- readRDS('dPrune-01-19-23.rds')

@@

### dplyr RMSE vs data.table RMSE
```{r}
dpVsDt=microbenchmark(
  dplyrMethod={
  fitVaried <- tv %>% filter(condit=="Varied") %>% wrap_optim(.,wm,lossFun=RMSE.tb);
  fitConstant <- tv %>% filter(condit=="Constant") %>% wrap_optim(.,wm,lossFun=RMSE.tb)},
  dtMethod={
  fitVaried2 <- tv %>% filter(condit=="Varied") %>% wrap_optim(.,wm,lossFun=RMSE.tb2);
  fitConstant2 <- tv %>% filter(condit=="Constant") %>% wrap_optim(.,wm,lossFun=RMSE.tb2)},
  times=5
)
knitr::kable(summary(dpVsDt),format="markdown")
```
|expr        |       min|        lq|      mean|    median|        uq|       max| neval|cld |
|:-----------|---------:|---------:|---------:|---------:|---------:|---------:|-----:|:---|
|dplyrMethod | 11.282291| 12.996857| 13.558623| 13.970367| 14.658515| 14.885086|     5|a   |
|dtMethod    |  4.933999|  5.306232|  5.681235|  5.515397|  6.131193|  6.519355|     5|b   |

@@

### Separate fits, vs. nesting method vs. split method
```{r}
nestSplit<-microbenchmark(
separate={fitVaried <- tv %>% filter(condit=="Varied") %>% wrap_optim(.,wm,lossFun=RMSE.tb2);
fitConstant <- tv %>% filter(condit=="Constant") %>% wrap_optim(.,wm,lossFun=RMSE.tb2) },
nestGroups = tv %>% group_by(condit) %>% nest() %>% mutate(fit=map(data,~wrap_optim(.,wm,RMSE.tb2))),
splitGroups = tv %>% split(.$condit) %>% map(~wrap_optim(.,wm,RMSE.tb2)),
times=5
)
knitr::kable(summary(nestSplit),format="markdown") # 03/02/23 - Mac Pro
```
|expr        |      min|       lq|     mean|   median|       uq|      max| neval|cld |
|:-----------|--------:|--------:|--------:|--------:|--------:|--------:|-----:|:---|
|separate    | 6.394459| 6.483235| 6.709882| 6.508353| 7.059420| 7.103941|     5|a   |
|nestGroups  | 6.253850| 6.457956| 6.789962| 6.756942| 7.044965| 7.436094|     5|a   |
|splitGroups | 6.117184| 6.455866| 6.605814| 6.461572| 6.640798| 7.353648|     5|a   |

@@

### Computing RMSE over Raw trials vs. RMSE of blocked training performance
```{r}
trialBlock<-microbenchmark(
  trialFit = tv %>% split(.$condit) %>% map(~wrap_optim(.,wm,RMSE)),
  blockFit = tv %>% split(.$condit) %>% map(~wrap_optim(.,wm,RMSE.tb2)),
  times=5
)
knitr::kable(summary(trialBlock),format="markdown")
```

|expr     |      min|       lq|     mean|   median|       uq|      max| neval|cld |
|:--------|--------:|--------:|--------:|--------:|--------:|--------:|-----:|:---|
|trialFit | 2.665184| 2.674186| 2.806476| 2.803491| 2.843806| 3.045712|     5|a   |
|blockFit | 5.169167| 5.264218| 5.350769| 5.285151| 5.491828| 5.543483|     5|b   |

>>>>

---
title: Benchmarking 
date: last-modified
categories: [Simulation, ALM, R]
code-fold: show
code-tools: true
execute: 
  warning: false
  eval: false
editor: 
  markdown: 
    wrap: 72
---

```{r}

pacman::p_load(tidyverse,foreach,doParallel,future,furrr,here)
purrr::walk(c(here("Functions/alm_functions.R","Functions/Display_Functions.R")),source)

n_cores <- parallel::detectCores()

param_grid <- tibble(crossing(
  c = seq(.5,5,.25),
  lr = seq(0.01, 1,.1),
  noise_sd = c(0,.0001,0.001, 0.01),
  inNodes = c(5, 7,14,28),
  outNodes = c(16, 32,64)
))
nrow(param_grid)

gen_train <- function(trainVec=c(5,6,7),trainRep=3,noise=0){
  bandVec=c(0,100,350,600,800,1000,1200)
  if(class(trainVec)=="list"){trainVec=unlist(trainVec)}
  ts <- rep(seq(1,length(trainVec)),trainRep)
  noiseVec=rnorm(length(ts),mean=0)*noise
  if(noise==0) {noiseVec=noiseVec*0}
  tibble(trial=seq(1,length(ts)),input=trainVec[ts],vx=bandVec[trainVec[ts]]+noiseVec)
}
fit_alm <- function(data, c, lr, noise_sd, inNodes, outNodes) {
  mse_list <- replicate(5, {
    train_data <- data[, c("trial", "input", "cor")] %>% rename("vx" = cor)
    sim_result <- sim_train(
      dat = train_data,
      c = c,
      lr = lr,
      inNodes = inNodes,
      outNodes = outNodes,
      noise_sd = noise_sd
    )
    train_data$almTrain <- sim_result$almTrain
    mse <- mean((data$vx - train_data$almTrain)^2)
    mse
  })
  avg_mse <- mean(mse_list)
  return(avg_mse)
}

gt <- gen_train(trainVec=c(5,6,7),trainRep=8) %>% mutate(cor=vx,err=(800-0)*exp(-.1*seq(1,n()))+0,vx=cor-err)

furrr::furrr_options(seed = TRUE)
plan(multisession, workers = n_cores-1)
tff <- system.time({

param_grid <- param_grid %>% mutate(performance = future_map_dbl(seq_len(nrow(.)), function(idx) {
    fit_alm(gt, c = c[idx], lr = lr[idx], noise_sd = noise_sd[idx], inNodes = inNodes[idx], outNodes = outNodes[idx])
  },.options = furrr_options(seed = T)))
  best_paramsF <- param_grid %>%
    arrange((performance)) 
  bestF <- head(best_paramsF,1)
})

# cluster <- parallel::makeCluster(n_cores-1)                 
# doParallel::registerDoParallel(cluster)
# tfP <- system.time({
#   param_grid <- param_grid %>%
#     mutate(performance = foreach(idx = seq_len(nrow(.)), .combine = c) %dopar% {
#       fit_alm(gt, c = c[idx], lr = lr[idx], noise_sd = noise_sd[idx], inNodes = inNodes[idx], outNodes = outNodes[idx])
#     })
#   best_paramsP <- param_grid %>%
#     arrange((performance)) 
#   bestP <- head(best_paramsP,1)
#   })
# stopImplicitCluster()
#   tfP

tI <- system.time({
gt <- gen_train(trainVec=c(5,6,7),trainRep=8) %>% mutate(cor=vx,err=(600-0)*exp(-.1*seq(1,n()))+0,vx=cor-err)

param_grid <- param_grid %>%
  mutate(performance = map_dbl(seq_len(nrow(.)), ~ {
    fit_alm(gt, c = c[.x], lr = lr[.x], noise_sd = noise_sd[.x], inNodes = inNodes[.x], outNodes = outNodes[.x])
  }))
best_paramsI <- param_grid %>%
  arrange((performance)) 
bestI <- head(best_paramsI,1)
})

tI

cat(paste0("furr time=",round(tff["elapsed"],2)," \n ",
          # "foreach parallel time=",round(tfP["elapsed"],2)," \n ",
           " Standard Time=",round(tI["elapsed"],2)))

4.2 -6.2 times faster with furr on m1

| Runs | Machine   | Cores | Method   | Time (seconds) |
|------|-----------|-------|----------|----------------|
| 9120 | M1        | 9/10  | Furr     | 9.56           |
| 9120 | M1        | 9/10  | Standard | 58.8           |
| 912  | M1        | 8/10  | Standard | 5.7            |
| 912  | M1        | 8/10  | Parallel | 1.2            |
| 912  | M1        | 9/10  | Furr     | 1.3            |
| 912  | 2015 iMac | 2/4   | Standard | 26.2           |
| 912  | 2015 iMac | 2/4   | Parallel | 18.5           |
| 912  | 2015 iMac | 3/4   | Furr     | 12.88          |
| 912  | GTX       | 3/4   | Standard | 25.32          |
| 912  | GTX       | 3/4   | Furr     | 13.1           |

------------------------------------------------------------------------

Furrr and standard version both work equally fast, tested with up to 120
simulation repetitions

```{r}
library(furrr)
furrr::furrr_options(seed = TRUE)
plan(multisession, workers = parallel::detectCores())

@@

d <- readRDS(here('dPrune-01-19-23.rds'))

@@

### dplyr RMSE vs data.table RMSE

```{r}
dpVsDt=microbenchmark(
  dplyrMethod={
  fitVaried <- tv %>% filter(condit=="Varied") %>% wrap_optim(.,wm,lossFun=RMSE.tb);
  fitConstant <- tv %>% filter(condit=="Constant") %>% wrap_optim(.,wm,lossFun=RMSE.tb)},
  dtMethod={
  fitVaried2 <- tv %>% filter(condit=="Varied") %>% wrap_optim(.,wm,lossFun=RMSE.tb2);
  fitConstant2 <- tv %>% filter(condit=="Constant") %>% wrap_optim(.,wm,lossFun=RMSE.tb2)},
  times=5
)
knitr::kable(summary(dpVsDt),format="markdown")
```

| expr        |       min |        lq |      mean |    median |        uq |       max | neval | cld |
|:-------|-------:|-------:|-------:|-------:|-------:|-------:|-------:|:-------|
| dplyrMethod | 11.282291 | 12.996857 | 13.558623 | 13.970367 | 14.658515 | 14.885086 |     5 | a   |
| dtMethod    |  4.933999 |  5.306232 |  5.681235 |  5.515397 |  6.131193 |  6.519355 |     5 | b   |

@@

### Separate fits, vs. nesting method vs. split method

```{r}
nestSplit<-microbenchmark(
separate={fitVaried <- tv %>% filter(condit=="Varied") %>% wrap_optim(.,wm,lossFun=RMSE.tb2);
fitConstant <- tv %>% filter(condit=="Constant") %>% wrap_optim(.,wm,lossFun=RMSE.tb2) },
nestGroups = tv %>% group_by(condit) %>% nest() %>% mutate(fit=map(data,~wrap_optim(.,wm,RMSE.tb2))),
splitGroups = tv %>% split(.$condit) %>% map(~wrap_optim(.,wm,RMSE.tb2)),
times=5
)
knitr::kable(summary(nestSplit),format="markdown") # 03/02/23 - Mac Pro
```

| expr        |      min |       lq |     mean |   median |       uq |      max | neval | cld |
|:-------|-------:|-------:|-------:|-------:|-------:|-------:|-------:|:-------|
| separate    | 6.394459 | 6.483235 | 6.709882 | 6.508353 | 7.059420 | 7.103941 |     5 | a   |
| nestGroups  | 6.253850 | 6.457956 | 6.789962 | 6.756942 | 7.044965 | 7.436094 |     5 | a   |
| splitGroups | 6.117184 | 6.455866 | 6.605814 | 6.461572 | 6.640798 | 7.353648 |     5 | a   |

@@

### Computing RMSE over Raw trials vs. RMSE of blocked training performance

```{r}
trialBlock<-microbenchmark(
  trialFit = tv %>% split(.$condit) %>% map(~wrap_optim(.,wm,RMSE)),
  blockFit = tv %>% split(.$condit) %>% map(~wrap_optim(.,wm,RMSE.tb2)),
  times=5
)
knitr::kable(summary(trialBlock),format="markdown")
```

| expr     |      min |       lq |     mean |   median |       uq |      max | neval | cld |
|:-------|-------:|-------:|-------:|-------:|-------:|-------:|-------:|:-------|
| trialFit | 2.665184 | 2.674186 | 2.806476 | 2.803491 | 2.843806 | 3.045712 |     5 | a   |
| blockFit | 5.169167 | 5.264218 | 5.350769 | 5.285151 | 5.491828 | 5.543483 |     5 | b   |

### 2023-05-23 Model/abc_htw.qmd
more image changes and misc folder

---
title: Approximate Bayesian Fitting
date: last-modified
categories: [Modeling, ALM, R, Bayesian]
image: images/abc_varied.png
page-layout: full
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

@@

ggsave("images/abc_constant.png",postC/(gcp + gco),width=12,height=10)

>>>>

---
title: Approximate Bayesian Fitting
date: last-modified
categories: [Modeling, ALM, R, Bayesian]
image: ../images/abc_varied.png
page-layout: full
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

@@

ggsave(here("images/abc_constant.png",postC/(gcp + gco),width=12,height=10))
